---
title: "Vector field of seal movements"
author: "Chiara Todaro"
date: "3 April 2019"
output:
  ioslides_presentation: default
  beamer_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r libraries,message=FALSE}
rm(list = ls())
library(plotly);library(ggplot2);library(dplyr)
```

## Data 
The **seals** data set from the *ggplot2* R package consists of

```{r data,collapse=TRUE}
data(seals)
summary(seals);

delta_s<-with(seals,cbind(delta_lat,delta_long));

sealsT<-mutate(seals,end_lat=lat+delta_lat,end_long=long+delta_long, #deltas=delta_s,
              displacement=apply(delta_s,1,norm,type="2"),
              sign_dlat=sign(delta_lat),sign_dlong=sign(delta_long))

```

It represents the vector field of seal movements in the Pacific Ocean.

* The **long** and **lat** variables represent the x and y coordinates of a grid where the 
measurements are taken. 
* The **delta_long** and **delta_lat** variables represent the x and y components of the displacement vector.

## Vector field

```{r vector_field,message=FALSE,warning=FALSE,fig.height=6.5,fig.width=6.5,fig.align="center"}
geo1 <- list(scope = 'north america', projection = list(type = 'natural earth'),
   showland = TRUE,landcolor = toRGB("darkolivegreen4"))
p1 <- plot_geo(locationmode = 'USA-states') %>%
   add_markers(data = sealsT, x= ~end_long, y = ~end_lat, #type = 'scatter',
               marker=list(symbol=I(4),alpha = 0.1,color="blue", size=3)) %>% #
   add_segments(data =sealsT,x = ~long, xend = ~end_long,
                y = ~lat, yend = ~end_lat, color=I("blue"),alpha=1) %>%
      #  size = I(1) ,alpha = 0.3
   layout(title = 'Vector field of seal movements',geo = geo1, showlegend = FALSE)
p1
```


## Contour plot



<div class="columns-2">

```{r contour_plot,message=FALSE,warning=FALSE,fig.height=4.0,fig.width=4.5,fig.align="center"}
p3<-plot_ly(data=sealsT, x = ~long, y = ~lat,
   z =~displacement, #zeta,#colorscale="Jet",
   type = "contour", contours = list(showlabels = TRUE)) 

p3
```


The grid had 21 points for the latitude and 55 points for the longitude.

The magnitude of the displacement is color-coded on the longitude-latitude grid 
with added contour lines.

Overall the displacement increases in the south-east direction.


</div>


## Radar plot


```{r lat_long_data}
summ_lat<-sealsT %>% dplyr::group_by(lat) %>%
   summarise(displMean=mean(displacement),deltalatMean=mean(delta_lat),deltalongMean=mean(delta_long))

summ_long<-sealsT %>% dplyr::group_by(long) %>%
   summarise(displMean=mean(displacement),deltalatMean=mean(delta_lat),deltalongMean=mean(delta_long))

legendtitle <- list(yref='paper',xref="paper",y=1.05,x=1.25,
                    text="mean lat/long displ",showarrow=F) #

p2 <- plot_ly(type = 'scatterpolar',mode="markers+lines") %>% #, size = 1,alpha = 0.8
   add_trace(r = summ_lat$displMean, #,summ_seals$deltarMean[Nseals]
            theta = summ_lat$lat,name = 'latitude',
            color=summ_lat$deltalatMean,fill = 'toself') %>%#,colorscale="Jet" #,"a"
   add_trace(r = summ_long$displMean,
            theta =  summ_long$long,name = 'longitude',
            color=summ_long$deltalongMean,fill = 'toself') %>% #,color = "green"
   layout(showlegend = TRUE, annotations=legendtitle,
          polar = list(radialaxis = list(visible = T)))


```

<div class="columns-2">

```{r radar_plot,message=FALSE,warning=FALSE,fig.height=5.5,fig.width=4.0}
p2
```

Mean displacement for fixed longitude and latitude. 
Colors code the mean longitude and latitude components of the mean displacement.


Notice from the sign of the color that the movement go from 45 to 30 
degrees of latitude, and from -120 to -180 degrees of longitude. 
On the latitude the displacement magnitude increase linearly, while 
there is an increase/decrease around -135 degrees on the longitude.

</div>



